Comb associative operads

### 2018-07-09 Magmatic_operads.tex
Modification algo de Buchberger

\section{Magmatic operads}

@@

Given an operad $\Oca$, we denote by $H_{\Oca}$ the \Def{Hiblert serie} of $\Oca$, that is
\begin{equation} \label{equ:definition_of_Hilbert_series}
  H_{\Oca}(t)=\sum_{n\in\N}\dim{\left(\Oca(n)\right)}t^n.
\end{equation}

@@

\begin{Theorem} \label{thm:Grassmann_formula_for__Hilbert_series_of_QMag}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. We have
  \begin{equation} \label{equ:lattice_structure_and_Hiblert_series}
    H_{\Oca_1\wedge\Oca_2}(t)+H_{\Oca_1\vee\Oca_2}(t)=H_{\Oca_1}(t)+H_{\Oca_2}(t).
    \end{equation}
  \end{Theorem}

\begin{proof}
  We have
  \[\begin{split}
  H_{\Oca_1\wedge\Oca_2}(t)+H_{\Oca_1\vee\Oca_2}(t)&=\sum_{n\in\N}\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_1\vee\Oca_2(n)\right)}t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}+\dim{\left(\Oca_1\vee\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\dim{\left(\Oca_1(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_2(n)\right)}t^n\\
  &=H_{\Oca_1}(t)+H_{\Oca_2}(t).
  \end{split}\]
\end{proof}
The third equality is due to Lemma~\ref{lem:lattice_structure_and_coefficients_of_Hiblert_series}.
% Hilbert series, enumeration, a little of rewrite systems, etc.

>>>>

\section{Magmatic operads}
\label{sec:Magmatic_operads}

@@

@@

\begin{Theorem} \label{thm:Grassmann_formula_for__Hilbert_series_of_QMag}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. We have
  \begin{equation} \label{equ:lattice_structure_and_Hiblert_series}
    \HilbertSeries_{\Oca_1\wedge\Oca_2}(t)+\HilbertSeries_{\Oca_1\vee\Oca_2}(t)=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
    \end{equation}
  \end{Theorem}

\begin{proof}
  We have
  \[\begin{split}
  \HilbertSeries_{\Oca_1\wedge\Oca_2}(t)+\HilbertSeries_{\Oca_1\vee\Oca_2}(t)&=\sum_{n\in\N}\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_1\vee\Oca_2(n)\right)}t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}+\dim{\left(\Oca_1\vee\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\dim{\left(\Oca_1(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_2(n)\right)}t^n\\
  &=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
  \end{split}\]
\end{proof}
The third equality is due to Lemma~\ref{lem:lattice_structure_and_coefficients_of_Hiblert_series}.
% Hilbert series, enumeration, a little of rewrite systems, etc.

### 2018-07-09 Operads_rewrite.tex
Modification algo de Buchberger

We have provided here definitions about operads in the category of sets.
Nevertheless, operads can be defined in the category of $\K$-vector
spaces, where $\K$ denotes in all this text a field of characteristic
zero. We call them \Def{linear operads}. All the above definitions
extend for linear operads, mainly by substituting Cartesian products
$\times$ of sets with tensor products $\otimes$ of spaces, maps with
linear maps, operad congruences with operad ideals, and cardinalities
of sets with space dimensions (for instance
in~\eqref{equ:hilbert_series}). Besides, when $\Oca$ is a linear operad
admitting a basis $B$ such that its unit $\Unit$ belongs to $B$ and all
partial composition maps are internal in $B$, $\Oca$ can be studied as a
set-theoretic operad.
\medbreak

@@

Let $\Rew$ be a set of rewrite rules and $\RewContext$ be rewrite
relation induced by $\Rew$. Since $\RewContext$ is in particular a
binary relation on $\Mag$, the classical notations about closures apply
here: we denote by $\RewContextT$ (resp. $\RewContextRT$,
$\RewContextRST$) the transitive (resp. reflexive and transitive,
reflexive, symmetric, and transitive) closure of $\RewContext$.
\medbreak

@@

A \Def{normal form} for $\RewContext$ is a binary tree $\Tfr$ such
that for all binary trees $\Tfr'$, $\Tfr \RewContextRT \Tfr'$ imply
$\Tfr' = \Tfr$. In other words, a normal form for $\RewContext$ is a
tree which is not rewritable by $\RewContext$. The set of all the normal
forms of $\RewContext$ is denoted by $\NormalForms_{\RewContext}$. The
trees of $\NormalForms_{\RewContext}$ admit the following description,
useful for enumerative prospects.
\medbreak

@@

When $\Rew$ is a terminating but not convergent orientation of $\Congr$,
we shall use a variant of the \Def{Buchberger semi-algorithm} for
operads~\cite[Section 3.7]{DK10} to compute a set of rewrite rules
$\Rew'$ such that, as binary relations $\Rew \subseteq \Rew'$, and
$\Rew'$ is a convergent orientation of $\Congr$. This semi-algorithm
takes as input a finite set of rewrite rules $\Rew$ and outputs the set
of rewrite rules $\Rew'$ satisfying the property stated above. Here is,
step by step, a description of its execution:
\begin{enumerate}[label={(\it\arabic*)}]
    \item Set $\Rew' := \Rew$.
    \item \label{item:loop_start_buchberger}
    If there is no branching pairs for $\RewContext'$ of trees of
    degrees $k$ for all $0 \leq k \leq 2 \, \Deg(\Rew') - 1$, the
    execution stops and the output is $\Rew'$.
    \item Otherwise, let $\Tfr$ be the branching tree for $\RewContext'$
    with the minimal degree and the minimal prefix word.
    \item Let $\{\Sfr_1, \Sfr_2\}$ be the branching pair for $\Tfr$ such
    that $\Sfr_1$ and $\Sfr_2$ have the minimal prefix words.
    \item Add to $\Rew'$ the rewrite rule
    \begin{math}
        \max_\leq \left\{\Sfr_1, \Sfr_2\right\}
        \Rew'
        \min_\leq \left\{\Sfr_1, \Sfr_2\right\}.
    \end{math}
    \item Go to Step~\ref{item:loop_start_buchberger}.
\end{enumerate}
The set of rewrite rules $\Rew'$ outputted by this semi-algorithm is the
\Def{completion} of $\Rew$. By Lemma~\ref{lem:degree_confluence},
$\RewContext'$ is confluent. Notice that, for certain inputs $\Rew$,
this semi-algorithm never stops. Notice also that the computed
completion depend on the total order $\leq$ on the binary trees of a
same arity.
\medbreak

>>>>

We have provided here definitions about operads in the category of sets.
Nevertheless, operads can be defined in the category of $\K$-vector
spaces, where $\K$ denotes in all this text a field of characteristic
zero. We call them \Def{linear operads} and we use them in
Section~\ref{sec:Magmatic_operads}. All the above definitions extend for
linear operads, mainly by substituting Cartesian products $\times$ of
sets with tensor products $\otimes$ of spaces, maps with maps, operad
congruences with operad ideals, and cardinalities of sets with space
dimensions (for instance in~\eqref{equ:hilbert_series}). Besides, when
$\Oca$ is a linear operad admitting a basis $B$ such that its unit
$\Unit$ belongs to $B$ and all partial composition maps are internal in
$B$, $\Oca$ can be studied as a set-theoretic operad. 
\medbreak

@@

Let $\Rew$ be a set of rewrite rules and $\RewContext$ be rewrite
relation induced by $\Rew$. Since $\RewContext$ is in particular a
binary relation on $\Mag$, the classical notations about closures apply
here: we denote by $\RewContextT$ (resp. $\RewContextRT$,
$\RewContextRST$) the transitive (resp. reflexive and transitive,
reflexive, symmetric, and sitive) closure of $\RewContext$.
\medbreak

@@

A \Def{normal form} for $\RewContext$ is a binary tree $\Tfr$ such
that for all binary tree $\Tfr'$, $\Tfr \RewContextRT \Tfr'$ implies
$\Tfr' = \Tfr$. In other words, a normal form for $\RewContext$ is a
tree which is not rewritable by $\RewContext$. A normal form for
$\RewContext$ of a binary tree $\Tfr$ is a normal form $\overline{\Tfr}$
for $\RewContext$ such that $\Tfr\RewContextRT\overline{\Tfr}$. When no
confusion is possible, we simply say normal form instead of normal form
for $\RewContext$. The set of all the normal forms is denoted by
$\NormalForms_{\RewContext}$. The trees of $\NormalForms_{\RewContext}$
admit the following description, useful for enumerative prospects.
\medbreak

@@

When $\Rew$ is a terminating but not convergent orientation of $\Congr$,
we shall use a variant of the \Def{Buchberger semi-algorithm} for
operads~\cite[Section 3.7]{DK10} to compute a set of rewrite rules
$\Rew'$ such that, as binary relations $\Rew \subseteq \Rew'$, and
$\Rew'$ is a convergent orientation of $\Congr$. This semi-algorithm
takes as input a finite set of rewrite rules $\Rew$ and outputs the set
of rewrite rules $\Rew'$ satisfying the property stated above. Here is,
step by step, a description of its execution:
\begin{enumerate}[label={(\it\arabic*)}]
\item Set $\Rew' := \Rew$ and let $\BranchingTrees$ be the set of
  branching trees for $\RewContext$.
\item \label{item:loop_start_buchberger}
  If $\BranchingTrees$ is empty, the execution stops and the output
  is $\Rew'$.
\item Otherwise, let $\Tfr$ be the branching tree for $\RewContext'$
  with the minimal degree and the minimal prefix word and remove $\Tfr$
  from $\BranchingTrees$.
\item Let $\{\Sfr_1, \Sfr_2\}$ be the branching pair for $\Tfr$ such
  that $\Sfr_1$ and $\Sfr_2$ have the minimal prefix words.
\item Let $\overline{\Sfr_1}$ and $\overline{\Sfr_2}$ be normal forms of
  $\Sfr_1$ and $\Sfr_2$, respectively.
\item If $\overline{\Sfr_1}$ is different from $\overline{\Sfr_2}$, add
  to $\Rew'$ the rewrite rule
    \begin{math}
      \max_\leq \left\{\overline{\Sfr_1}, \overline{\Sfr_2}\right\}
      \Rew'
      \min_\leq \left\{\overline{\Sfr_1}, \overline{\Sfr_2}\right\}.
    \end{math}
  \item Go to Step~\ref{item:loop_start_buchberger}.
\end{enumerate}
The set of rewrite rules $\Rew'$ outputted by this semi-algorithm is the
\Def{completion} of $\Rew$. By Lemma~\ref{lem:degree_confluence},
$\RewContext'$ is confluent. Notice that, for certain inputs $\Rew$,
this semi-algorithm never stops. Notice also that the computed
completion depend on the total order $\leq$ on the binary trees of a
same arity.
\medbreak
